---
title: "Statistical Inference Project: A simulation exercise"
author: "Paul Reiners"
date: "January 18, 2015"
output: html_document
---

Exponential distribution CLT
----------------------------

Compare the distribution of 1000 exponentials

```{r}
lambda <- 0.2
n <- 1000
hist(rexp(n, lambda))
```

and the distribution of 1000 averages of 40 exponentials

```{r}
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(rexp(40, lambda)))
hist(mns)
```

Notice how the distribution of the 1000 **averages** of random exponentials appears to be approximately normal.  Also note that its mean appears to be the mean of the 1000 generated random exponentials.

Sample and theoretical means
----------------------------

What is the sample mean of the distribution?

```{r}
mean(mns)
```

The theoretical mean of the exponential distribution is 1/lambda:

```{r}
1/lambda
```

They differ by less than 1%:

```{r}
abs(mean(mns) - 1/lambda) / (1/lambda)
```

Sample and theoretical variances
--------------------------------

What is the sample variance of the distribution?

```{r}
vars = NULL
for (i in 1 : 1000) vars = c(vars, var(rexp(40, lambda)))
sample.var <- mean(vars)
sample.var
```

The theoretical standard deviation of the exponential distribution is 1/lambda.  So the variance is the square of this:

```{r}
theoretical.var <- (1/lambda)^2
theoretical.var
```

They also differ by less than 1%:

```{r}
abs(sample.var - theoretical.var) / theoretical.var
```
